TacticToe: Learning the Game of Tactics

### 2018-01-28 tactictoe_jv.tex
predictions: almost done

\todo{TG : comparison with hammers in usability}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight}, 
  \isabelle~\cite{isabelle} 
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant 
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer 
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

\todo{TG: rewrite the plan}
  To do this, we \emph{extract and record} tactic invocations from the ITP 
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a 
  guidance for a \emph{Monte Carlo tree search} that explores the different 
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance 
  of the single best \tactictoe strategy exceeds the performance of a hammer 
  system used with a single strategy and a single efficient external prover.

\subsection{The problem}
How a user works with an ITP and how this gives rise to a machine learning problem.

Two ways of representing a formulas.

Definition of a tactic in \holfour. In this research, we will take the definition
that a tactic takes a goal and returns a list of goals.

Input of the prediction: goal

@@

\todoi{add prediction of 4 kinds of things}
\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.

@@

\section{Prediction}\label{s:prediction}
The learning-based selection of relevant lemmas significantly improves the 
automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we adapt 
one of the strongest lemma selection 
methods: the modified distance-weighted \emph{$k$ nearest-neighbour} 
classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects: 
 goals, theorems and list of goals.

We first introduce
specificities associated with the prediction of each object: the dataset from
which objects are selected and the different use cases for the 
prediction. We then details how we predict objects suitable for the current 
goal by comparing their features with the feature of the goal.

\paragraph{Tactics}
Given a goal $g$, the classifier selects a list of input goals
ordered by their similarity to $g$. Since each input goal is associated 
with a tactic in our database, the ordered list of goals induces a ordered  
list of associated tactics which gives us a prior policy function for  MCTS. 
And tactics are tried in this order during the 
proof search. Therefore, to order the tactics correctly, it is crucial to 
estimate the similarity between goals with mathematically relevant features.

In all proofs, we record each tactic invocation and link (associate) the 
currently open goal with the tactic's name in our database.
Given a new open goal $g$, the \emph{score of a tactic $T$ wrt. $g$} is defined 
to be the score (similarity) of $T$'s 
associated goal which is most similar to $g$.
The idea is that tactics with high scores are more likely to solve the open 
goal $g$, 
since they were able to solve similar goals before.

Whenever we mention that we want to predict a goal, we 
actually intend to predict its associated tactics.
We derive from tactic selection a prior policy for 
our search algorithm. Lists of goals (products of tactics) evaluate the success 
of the policy and adapt it accordingly.

\paragraph{Theorems}
Because we may augment the dataset of tactics with abstracted tactics where 
list of theorems have been erased, we need to find a way to instantiate them.

This algorithm is also useful for selecting premise to ATPs. (see 
Section~\ref{sec:atp})

Given a open goal $g$, the classifier selects a list of theorems ordered by 
their 
similarity to $g$.
A dataset of theorems is constructed from the \holfour database and local 
theorem values.
We predict theorems most similar to the goal for instantiate list of theorems 
that appears as arguments of abstracted tactics (see Section 
\ref{sec:synthesis}). A list of 
predicted theorems 
are also given as input to ATPs. The number of predictions is 16 for tactic 
arguments and \metis, but is increased to 128 for \eprover.

\paragraph{List of goals}
The dataset of learned list of goals is compiled during orthogonalization.
This set is separated into positive and negative examples.
Given a list of goals outputted by a tactic,
we compute the closest lists of goals in our dataset. The ratio of positive 
examples from the total number of examples gives us a prior evaluation function 
for MCTS (see Section~{}).

@@

The first co-distance $sim_1$ adds a modified weight of each shared 
features to compute the total score.
In a second co-distance $sim_2$, we additionally take into account 
the total number of features to reduce the seemingly unfair advantage of big 
feature sets in the first scoring function.
\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{6}}}\]
\[sim_2 (f_o, f_p) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{6}}}}
{ln (e + card(f_o) + card(f_p))}
\]

Making prediction for an object $o$ consists sorting a set of objects 
by their similarity to $o$.

We use $sim_1$ for goals and theorems and $sim_2$ for list of goals.
The similarity $sim_1$ is used for predicting a tactic (via goals) because if a
tactic is effective on large goal $g$ then it is often suitable for a goal $g'$ 
containing sub-formulas of $g$. The same monotonicity heuristic can justify the 
use of $sim_1$ in theorems. Indeed, in practice, a large theorem is often a 
conjunction of small theorems from the same domain. And if a conjunction 
contains a formula related to a problem, then the other conjuncts 
from the same domain may also help.

@@

The predictions algorithm is called many times during proof search to chose 
suitable tactics, instantiate tactic arguments by terms and theorems and
estimate the provability of a list of goals.
In order to speed up the predictions, we preselect 500 features vectors of 
tactics, theorems and list of goals.

We first rank feature vectors by their distance to the conjecture. But better 
ranked vectors only contains symbols from 
the conjecture and may not be suitable for later stages of the proof where the 
goals may have derived significantly from the conjecture. To solve this 
problem, we reorder the list so that descendants (see 
Definition~\ref{def:recdep}) of vectors inherits the similarity of their 
parent.

\begin{definition}(Recursive dependencies of a feature 
vector)\label{def:recdep}
Let $\mathfrak{F}$ be a set of recorded feature vectors. 
The set of all recursive dependencies $D_\infty$ of a feature vector $f$ is 
inductively defined by:

\begin{align*}
D_0 &=_{def} \lbrace f \rbrace \\
D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists 
(\hat{t},\hat{g}) \in D_n.\ g \in \hat{t}(\hat{g}) \rbrace  \\
D_\infty &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
\end{align*}
where $\hat{t}(\hat{g})$ denotes the resulting set of goals produced by the 
application 
of the tactic $\hat{t}$ to the goal $\hat{g}$.
\end{definition}

We augment the ordered predictions by inserting after each predicted vector 
its descendants and removing lower ranked duplicates. 
%This contrast with 
%theorem predictions where we only select parents. 
 This phase is only applied 
during preselection, so is of no great consequence as long as we select enough 
theorems and tactics.

Since predicting efficiently is crucial during proof 
search, we preselect a number of tactics before the search.
We take the first 500 vectors of the augmented ordered list. Only members of 
this list can be predicted during proof search. When an 
open goal $g$ is created, the scores of those 500 preselected tactics are
recomputed which gives a preferential order of tactics to be tried on $g$.

\section{Orthogonalization}\todo{MN + RK: review + fix typos}

@@

We made argument predictors for one type of arguments: list 
of theorems. We choose this very common type because perfect instantiation of 
those arguments would lead to the largest improvements (see 
Section~\ref{sec:perfect_exp}). We plan to 
extend our algorithm to term which is another frequent argument types.
But term prediction have been so far proven to weak to positively 
influence the proof search.

@@

\begin{align*}
\mathbb{A}^{0} &=_{def} 
\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\ 
\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\ 
\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n\\
\mathbb{G}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
\end{align*}

By the base case of these definition, a goal is solved if there exists a tactic 
applied to this goal that produces no goals.
\end{definition}

\subsection{Prior tactic policy}
The estimation through the distance produced by the tactic selectors are hard 
to translate into a prior tactic policy, so we rely only on the order of 
tactics to estimate the probability with which a tactic should be tried.
Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
distance to a goal $g$. Let $c_{policy}$ be a constant heuristic that estimates 
how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
tactics $t_{i+1},\ldots,t_{500}$. We make the strong assumption that $c_policy$ 
is constant across the search. We can now calculate the prior policy by:

@@

\subsection{Prior node evaluation}

We now concentrate on defining a reasonable evaluation function that also 
influences node selection in MCTS. Essentially, we estimate how 
interesting is the list of goals of a tactic to determine if we should continue 
searching in this direction. Theoretically, an interesting list of goal is one 
that has the shortest proof, and we could measure the likelihood of being 
provable in a number of steps from previous proof search. However, this 
approach produces a lot of negatives\todo{TG: cite Hard Negative mining (cite 
Deep math)}. So to reduce this number, we limit 
ourselves to feature vectors tested during orthogonalization. We declare a list 
of goals l to be positive, noted $l \in Positive$, if it is has been 
produced by the winner of an orthogonalization competition. We extend the 
definition of a $distance$ to a list of goals by considering the union of the 
features of the goals. The set of the $r$ set of goals closest to a set of 
goal $l$ is noted $Neighbour(r,l)$. We can define the set of positives and 
negatives within a radius $r$ of a 
list of goals $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
Positive \rbrace$.

During our experiments, we evaluate output goals in three different manners 
with the following functions:

\begin{align*}
PriorEvaluation_{none} (a_i) &= 0 \\
PriorEvaluation_{productive} (a_i) &= 1\\
PriorEvaluation_r (a_i) &= \frac{card\ Positive(r,a_i)}{r}\\
\end{align*}

For all evaluations functions, if a tactic is not productive then there is no 
output to evaluate, so the result is a failure and a reward of $0$ will be 
given for this node. The same thing happens if the node selected is already 
proven or a descendant of a proven goal. 
Experimentally, it is preferable not to 
filter solved (or descendants of solved) nodes before 
node selection to keep the estimation of the algorithm balanced.

@@

@@

\section{Experimental Evaluation}\label{s:experiments}\todo{TG: introduction + 
consistent tables}

Disclaimer:
Conditions of the experiments may vary slightly between different tables 
because of various factors such 
as overhead cost of additional features or the fact that they were run on 
different machines on a local machine or on the server.

\subsection{Methodology} \todo{TG: is copied, update references of sections and
new algorithm + mention when orthogonalization is done}

The evaluation imitates the construction of the library: For each theorem only 
the previous human proofs are known. These are used as the learning base for 
the predictions.
To achieve this scenario we re-prove all theorems during a modified build of 
\holfour.
As theorems are proved, their tactical proofs and their statements are recorded 
and included in the 
training examples.
For each theorem we first attempt to run the \tactictoe search with a time 
limit of 5 seconds,
before processing the original proof script.
In this way, the fairness of the
experiments is guaranteed by construction. 
Only previously declared \sml 
variables (essentially tactics, theorems and simpsets) are accessible. 
And for each theorem to be re-proven \tactictoe is only trained on previous 
proofs.

\paragraph{Dataset}

7000+ top-level theorems that were created by . 
In this evaluation, we do not include the 250 theorems created by

All top-level theorems created by a call to 
\texttt{store\_thm}, 
\texttt{Q.store\_thm} or \texttt{maybe\_thm} are considered with the exception 
of 440 calls that contains a let binding. They are typically harder to re-prove.
We also ignore the 1175 \texttt{save\_thm} calls. Most of them are easily 
provable theorems using instantiations or conjuncts of previous 
theorems.

%\begin{itemize}
%%\item the , these are 
%%typically 
%%harder for \tactictoe as it may not have seen the particular local 
%%declaration 
%%before and therefore would have to find an alternative proofs. We also do not 
%%learn from these proofs because they were explictly local.
%%\item the 1175 calls to save\_thm. These are typically mostly trivial and 
%%easy 
%%theorems.
%%The remaining 7088 top-level theorems are evaluated in the full scale 
%%evaluation and the level of difficulty of these problem is assessed by a 
%%comparison with \eprover.
%\end{itemize}

Among these types of theorems, every tenth theorems of the first third of the 
library is evaluated during tuning, which give us a total of 273 theorems to 
train different parameters on.

\paragraph{Strategy selection}
Although the training process in each strategy on its own is fair, the 
selection of the 
best strategy in Section~\ref{sec:sel_param} should also be considered as a 
learning 
process. To ensure the global fairness, the final experiments in 
Section~\ref{sec:exp_full} 
runs the best strategy on the full dataset which is about 10 times larger. The
performance is minimally better on this validation set.

\subsection{Tuning \tactictoe}

\subsubsection{Prediction parameters}\todo{TG: remove this section or mention
that parameters were optimized in the previous version}
The first experiment concerns the choice of the right kind of features and 
feature scoring 
mechanism. The results are presented in Table~\ref{tab:feature_param}.
We observe that the higher-order features and the feature of the top logical 
structure increase minimally the number of problems solved. The attempted 
length penalty on the total number of features is actually slightly harmful.
We did not turn on the top feature flag in the next experiments because they 
are slightly more expensive too compute.

%
%\begin{table}[t]
%\centering\ra{1.3}
%\small
%\begin{tabular}{llll}
%\toprule
% ID & Learning parameter & Solved & $U(D_1)$ \\
%\midrule
% $D_0$ & $codist_1$ (length penalty) & 172 (20.0\%) & 5 \\
% $D_1$ & $codist_0$ (default) & 179 (20.8\%) & 0 \\
% $D_2$ & no top features & 175 (20.3\%) & 18 \\
% $D_3$ & no higher-order features & 178 (20.7\%) & 8 \\
%\bottomrule
%\end{tabular}
%\caption{\label{tab:feature_param}Success rate of strategies with different 
%learning parameters on the training set.}
%\end{table}
%
%\subsubsection{Search parameters}
%
%\paragraph{tactic timeout}
%\begin{table}[t]
%\centering\ra{1.3}
%\small
%\begin{tabular}{llll}
%\toprule
% ID & Searching parameter & Solved & $U(D_{9})$ \\
%\midrule
% $D_1$ & $codist_0$ (default)      & 179 (20.8\%) & 10 \\
% $D_4$ & tactic timeout 0.004 sec & 175 (20.3\%) & 9 \\
% $D_5$ & tactic timeout 0.1 sec   & 178 (20.7\%) & 10 \\
%\bottomrule
%\end{tabular}
% \caption{\label{tab:search_param}Success rate of strategies with different 
% search parameters on the training set.}
%\end{table}

\subsubsection{MCTS parameters: policy and evaluation}

Width did include failing tactics in this experiments. $c_{exploration}$ = 1.
$E_{none}$.

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{ll}
\toprule
 $c_{policy}$ & Solved \\
\midrule
 $2/3$ & 120\\
 $1/2$ & 125\\
 $1/3$ & 117\\
% 1/2 & Productive & No & $E_{none}$ & 1 & 120\\
% 1/2 & Productive & No & $E_{productive}$ & 1 & 127\\
\bottomrule
\end{tabular}
\caption{\label{tab:cost_param} 
%on a local machine
}
\end{table}

%This experiments were performed without the square root in the policy 
%function.

This experiments are performed with orthogonalization on and policy coefficient 
of $1/2$.

>>>>

\todo{Comparison with hammers in usability}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight}, 
  \isabelle~\cite{isabelle} 
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant 
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer 
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

\todo{Rewrite the plan}
  To do this, we \emph{extract and record} tactic invocations from the ITP 
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a 
  guidance for a \emph{Monte Carlo tree search} that explores the different 
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance 
  of the single best \tactictoe strategy exceeds the performance of a hammer 
  system used with a single strategy and a single efficient external prover.

\subsection{The problem}
How a user works with an ITP and how this gives rise to a machine learning 
problem.
Two ways of representing a formulas.
Definition of a tactic in \holfour. In this research, we will take the definition
that a tactic takes a goal and returns a list of goals.
Input of the prediction: goal
If a tactic produces no goals then its input goal is solved.

@@

\todoi{add prediction of 3 kinds of things}
\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.

@@

\section{Prediction}\label{s:prediction}
The learning-based selection of relevant lemmas significantly improves the 
automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we adapt 
one of the strongest lemma selection 
methods: the modified distance-weighted \emph{$k$ nearest-neighbour} 
classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects: 
 tactics, theorems and list of goals.

We first introduce specificities associated with the prediction of each object. 
In particular, we present the dataset from which objects are selected and the 
purpose of the selection process. 
We then explain the construction of the similarity measure by which relevant 
objects are predicted.

\paragraph{Tactics}
To start with, we build a database of tactics consisting
of (tactic,goal) pairs recorded from successful human proofs (see 
Section~\ref{sec:recording}).
Then, during proof search, we order recorded goals according to their 
similarity with a targeted goal $g$. This induces through the recorded pairs an 
order on tactics. The intuition is that tactics that have been successful on 
goals similar $g$ should be tried first. 
This predicted tactic order is translated into a prior policy for our MCTS 
algorithm (see Section~\ref{sec:policy}).
Additionally, tactic selection allows us to improve the quality of the database 
of tactics during orthogonalization (see Section~\ref{sec:ortho}).

\paragraph{Theorems as Argument of Tactics}
We first collect a set of theorems from the \holfour theorem database and local 
theorem values, for our theorem predictor to select from.
Predicted tactics during proof search and orthogonalization may contain 
abstracted tactics where list of theorems have been erased (see Section 
\ref{sec:synthesis}).
We instantiate those tactics by theorems that are syntactically the closest to 
the targeted goal as they are more likely to help.
The same algorithm selects suitable premises for ATPs integrated with  
\tactictoe (see Section~\ref{sec:atp}).

\paragraph{List of Goals as Output of Tactics}
A dataset of list of goals is compiled during orthogonalization (see 
Section~\ref{sec:evaluation}).
This set is separated into positive and negative examples.
List of goals that subsumes the result of the original tactic on a 
targeted goal are considered positive. During proof search, given a list of 
goals $l$ created by a tactic, we select lists of goals that are most similar 
to $l$.
The ratio of positive examples in the selection gives us a prior node 
evaluation for MCTS.

@@

The first co-distance $sim_1$ adds a modified weight of each shared 
features to compute the total score.
In a second co-distance $sim_2$, we additionally take into account 
the total number of features to reduce the seemingly unfair advantage of big 
feature sets in the first scoring function.
\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{6}}}\]
\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{6}}}}
{ln (e + card(f_o) + card(f_p))}\]

In our setting, making prediction for an object $o$ consists of sorting a set 
of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems 
and $sim_2$ for list of goals. The similarity $sim_1$ is used for predicting a 
tactic via its associated goal because if a
tactic is effective on large goal $g$ then it is often suitable for a goal $g'$ 
containing sub-formulas of $g$. The same monotonicity heuristic can justify the 
use of $sim_1$ in theorems. Indeed in practice, a large theorem is often a 
conjunction of small theorems from the same domain. And if a conjunction 
contains a formula related to a problem, then the other conjuncts 
from the same domain may also help.

@@

In order to speed up the predictions during the search for a proof of a 
conjecture $c$, we preselect 500 tactic-goal pairs and theorems. 
Preselection for list of goals is induced by the preselection of the input 
goals in the tactic-goal pairs. Those will 
be the only objects available to \tactictoe for proving $c$.

The first idea is to select tactic-goal pairs and theorems by their distance to 
the conjecture. However, this selection may not be adapted to later stages of 
the proof where goals may have derived significantly from the 
initial conjecture $c$. In order to 
anticipate the production of diverse goals, our prediction during preselection 
take into account dependencies between objects of the same dataset.
The dependence relation is assymetric. Through the relation, each object has 
multiple children and at most one parent.
Once a dependency relation is established, we can calculate a dependency score 
for each object which is the maximum of its similarity score and the similarity 
score of its parent. In the end, the 500 feature vectors in each dataset with 
the highest dependency score are preselected.

In the following, we give a mathematical definition for the dependency relation 
between labels for tactic-goal pairs and theorems.

\begin{definition}(Dependencies of tactic-goal pairs)\\
Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
The dependencies $D_\infty$ of a tactic-goal pair $(t_0,g_0)$ is 
inductively defined by:

\begin{align*}
D_0 &=_{def} \lbrace (t_0,g_0) \rbrace \\
D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists 
(\hat{t},\hat{g}) \in D_n.\ g \in \hat{t}(\hat{g}) \rbrace  \\
D_\infty &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
\end{align*}
where $\hat{t}(\hat{g})$ denotes the resulting set of goals produced by the 
application 
of the tactic $\hat{t}$ to the goal $\hat{g}$.
\end{definition}

\begin{definition}(Dependencies of a theorem)\\
Let $\mathfrak{T}$ be the set of recorded theorems.
The dependencies $D_\infty$ of a theorem $t$ are the set of theorems in 
$\mathfrak{T}$ appearing in the proof $t$ as recorded in \cite{hh4h4}.
Contrarily to the dependency relation in tactic-goal pairs, this relation is 
not a transitive closure.
\end{definition}

To preselect a list of goal $l$, we consider the tactic input $g$ that was 
at the origin of the production of this list. If and only if there exists a 
tactic $t$ such as $(t,g)$ is in the preselected tactic-goal pairs, the 
list of goals $l$ is preselected. This has the desirable effect that
negative and positive examples created during one orthogonalization 
step are chosen or discarded together. \todo{problematic forward reference?}

\section{Orthogonalization}\label{sec:ortho}

@@

We made argument predictors for one type of arguments: list 
of theorems. We choose it is the most common type and the structure of theorems 
are complex so they contain enough information to be predicted accurately. 
We plan to extend our algorithm to term which is another frequent argument 
types but term synthesis is so far too weak to positively 
influence the proof search.

@@

\begin{align*}
\mathbb{A}^{0} &=_{def} 
\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\ 
\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\ 
\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
\mathbb{G}^\infty =_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
\end{align*}
\end{definition}

\subsection{Prior tactic policy}\label{sec:policy}
Distances between goals produced by the tactic predictor are hard 
to translate into a prior tactic policy, so we rely only on the order of 
tactics to estimate the probability with which each tactic should be tried.
Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
distance to a goal $g$. Let $c_{policy}$ be a constant heuristic that estimates 
how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
tactics $t_{i+1},\ldots,t_{500}$. We make the strong assumption that $c_policy$ 
is constant across the search. We can now calculate the prior policy by:

@@

\subsection{Prior node evaluation}\label{sec:evaluation}

We now concentrate on defining a reasonable evaluation function that also 
influences node selection in MCTS. Essentially, we estimate how 
interesting is a list of goals produced by a tactic. This way, we can determine 
if we should continue searching in this direction. Theoretically, an 
interesting list of goal is one 
that has the shortest proof, and we could measure the likelihood of being 
provable in a number of steps from previous proof search. However, this 
approach produces a lot of negatives\todo{TG: cite Hard Negative mining (cite 
Deep math)}. So to reduce this number, we limit 
ourselves to tactics tested during orthogonalization. We declare a list 
of goals l to be positive, noted $l \in Positive$, if it is has been 
produced by the winner of an orthogonalization competition. The set of the $r$ 
goals closest to a list of goals $l$ is noted $Neighbour(r,l)$. We can 
define the set of positives within a 
radius $r$ of $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
Positive \rbrace$.

During our experiments, we evaluate output goals with the following function:

\begin{align*}
PriorEvaluation_r (a_i) &= \frac{card\ Positive(r,a_i)}{r}\\
\end{align*}

If a tactic is not productive then there is no output to evaluate, so the 
result is a failure and a reward of $0$ will be 
given for this node. Failure also happens if the node selected is already 
a proven goal or a descendant of a proven goal. Experimentally, it is 
preferable not 
to filter solved nodes and descendants of solved nodes before node selection 
to keep the estimations of the algorithm balanced.

@@

mention theorem dependencies for eprover but not for metis.

@@

\section{Experimental Evaluation}\label{s:experiments}
\todo{Josef: give the specification} 
Each re-proving experiment is performed on a single CPU of a server in Prague 
with the exception of the experiments 
relying on asynchronous \eprover calls that run on two CPUs.

\subsection{Methodology} 
The evaluation imitates the construction of the library: For each theorem only 
the previous human proofs are known. These are used as the learning base for 
the predictions.
To achieve this scenario we re-prove all theorems during a modified build of 
\holfour.
As theorems are proved, their tactical proofs and their statements are 
recorded and included in the training examples.
For each theorem we first attempt to run the \tactictoe search with a time 
limit of 60 seconds, before processing the original proof script.
In this way, the fairness of the experiments is guaranteed by construction. 
Only previously declared \sml 
variables (essentially tactics, theorems and simpsets) are accessible. 
And for each theorem to be re-proven \tactictoe is only trained on previous 
proofs.

\paragraph{Datasets: optimization and validation}\todo{rephrase to something 
more 
coherent}
All top-level theorems from the standard library are considered with the 
exception of 440 hard problems (containing a \texttt{let} construction in their 
proof) and 1175 easy problems (build from \texttt{save\_thm} calls). 
Therefore, during the full-scale experiments, we evaluate 7164 theorems.
We use every tenth theorem of the first third of the standard library of 
them for parameter optimization, which amounts to 273 theorems.
Although the evaluation of each set of parameters on its own is fair, 
the selection of the best strategy in Section~\ref{sec:tuning} should also be 
considered as a learning process. To ensure the global fairness, the final 
experiments in Section~\ref{sec:full_exp} runs the best strategy on the full 
dataset which is about 30 times larger.

\subsection{Tuning \tactictoe}~\label{sec:tuning}

Fix parameters: not optimized here. 
 - Features extraction method.
 - k-nn distance.
 - MCTS evaluation radius is set to 10. 
 - Metis : 16 lemmas.
 - Eprover: 64 lemmas and 5 seconds timeout 1 asynchronous call

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{llc}
\toprule
 Technique & Parameters & Solved \\
\midrule
Tactic time out & 0.02s & 155 (56.8\%)\\
                & 0.05s (default) & 155 (56.8\%)\\
                & 0.1s & 153 (56.0\%)\\
\midrule
Orthogonalization & $radius = 20$ (default) & 155 (56.8\%)\\ 
                  & $radius = 10$ & 125 (50\%)\\
                  & $radius = 40$ & 125 (50\%)\\
\midrule
Abstraction       & none (default) & 125 (50\%)\\
                  & $theorems = 8$ & 125 (50\%)\\
                  & $theorems = 16$ & 125 (50\%)\\
                  & $theorems = 32$ & 125 (50\%)\\
\midrule
MCTS Policy & $c_{policy} = 1/2$ (default) & 125 (50\%)\\
            & $c_{policy} = 2/3$ & 125 (50\%)\\
            & $c_{policy} = 1/3$ & 125 (50\%)\\
\midrule
MCTS Evaluation & none (default) & 125 (50\%) \\
           & $c_{exploration} =\sqrt{2}$ & 125 (50\%)\\ 
           & $c_{exploration} = 1$ & 125 (50\%)\\ 
           & $c_{exploration} = 2$ & 125 (50\%)\\ 
\midrule
ATP integration & none (default) & 125 (50\%)\\
                & \metis 0.1s & 125 (50\%)\\
                & \metis 0.2s & 125 (50\%)\\
                & \metis 0.4s & 125 (50\%)\\ 
                & \eprover & 125 (50\%)\\
                & \metis 0.1s + \eprover & 190\\
\midrule
All & best parameters (without \eprover) & \\ 
\bottomrule
\end{tabular}
\end{table}

%\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
%
%We investigate what are the limits of \tactictoe and what benefits additional 
%knowledge could provide. To this end, we will modify the set of features to 
%include features vectors that should not be known in a regular experiment. We  
%include some vectors from the recorded human proof and perform the proof 
%search 
%after the recording instead of before in all other experiments. This 
%experiment 
%is not fair and should only be compared to 
%experiments in the same settings.
%
%For clarity, we will name the set of feature vectors that can be recorded from 
%the current human proof $\mathbb{H}$ and the set of vectors recorded from 
%previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
%$\pi_1^1$ the projection on the 
%first component returning the label of a feature vector. And we define the 
%function $T$ that takes a tactic and return the list of its tokens.
%
%We will progressively add bigger and bigger 
%subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
%search success rates. 
%
%Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
%consider:
%\begin{align*}
%\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
%   \rbrace \\
%\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
%\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
%\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
%\rbrace\\
%\end{align*}
%
%
%$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
%prediction for those tactics.
%
%
%\begin{table}[h]
%\centering\ra{1.3}
%\small
%\begin{tabular}{lll}
%\toprule
% & Additional features & Solved \\
%\midrule
%$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
